### 2005-09-03 paper/abstract.tex
I like $\mathcal{O}$ for big-O notation

The problem we are addressing is the one of text indexing for approximate matching. We consider that we are given a text $\mathcal{T}$ which undergoes some preprocessing to generate an index. We can later query this index to identify the places where a string occurs up to a certain number of allowed substitutions $k$. We present a structure for this indexing which occupies space $\mathcal{O}(n\log^kn)$ in the average case, independent of alphabet size, $n$ being the text size. This structure support searching in $\mathcal{O}(m^{k+1})$ time, for patterns of size $m$, again independently of alphabet size. The construction of the structure has time bounded by $\mathcal{O}(N|\Sigma|)$, where $N$ is the number of nodes in the index and $|\Sigma|$ the alphabet size.