### 2005-09-16 paper/abstract.tex
Many small changes and it's almost done

The problem we address is text indexing for approximate matching. We consider that we are given a text $\mathcal{T}$ which undergoes some preprocessing to generate an index. We can later query this index to identify the places where a string occurs up to a certain number of allowed substitutions $k$. We present a structure for indexing which occupies space $\mathcal{O}(n\log^kn)$ in the average case, independent of alphabet size, $n$ being the text size. This structure support searching in $\mathcal{O}(m^{k+1})$ time, for patterns of size $m$, again independent of alphabet size. The construction of the structure has time bound by $\mathcal{O}(kN|\Sigma|)$, where $N$ is the number of nodes in the index and $|\Sigma|$ the alphabet size.

\textbf{Keywords:} string algorithms, suffix trees, approximate text matching, text indexing.