### 2005-09-19 paper/main.tex

\documentclass[11pt]{article}
\usepackage[letterpaper,hmargin=1in,vmargin=1in]{geometry}
\usepackage[vlined,linesnumbered,ruled]{algorithm2e}
\input{figures/transfig}
\usepackage[]{graphicx}
\usepackage{amssymb}
\usepackage{subfigure}
\usepackage{QED}

\newcommand*{\Assign}{\ensuremath\longleftarrow}
\newcommand{\putstring}[1]{\textsl{#1}}
\newtheorem{definition}{Definition}
\newtheorem{lemma}{Lemma}
\graphicspath{{figures/}{graphics/}}

\title{Dotted Suffix Trees\\A Structure for Approximate Text Indexing}
\author{Lu\'{\i}s Pedro Coelho \and Arlindo L. Oliveira}
\date{}

\begin{document}
\maketitle

\begin{abstract}
\input{abstract}
\end{abstract}

\section{Introduction}\label{sec:introduction}
\input{introduction}

\section{The Indexing Structure}\label{sec:structure}
\input{structure}

\section{Searching}\label{sec:search}
\input{search}

\section{Constructing the Dotted Tree}\label{sec:construction}
\input{construction}

\section{Space Considerations}\label{sec:space}
\input{space}

\section{Experimental Results}\label{sec:experimental}
\input{experimental}

\section{Conclusions}\label{sec:conclusion}
\input{conclusion}

\bibliographystyle{alpha}
\bibliography{biblio}
\end{document}