### 2006-07-04 paper/abstract.tex
edition distance

The problem we address is text indexing for approximate matching. We consider that we are given a text $\mathcal{T}$ which undergoes some preprocessing to generate an index. We can later query this index to identify the places where a string occurs up to a certain number of allowed errors~$k$ (edition distance). We present a structure for indexing which occupies space $\mathcal{O}(n\log^kn)$ in the average case, independent of alphabet size, $n$ being the text size. This structure can be used to report the existence of a match with $k$ errors in $\mathcal{O}(3^k m^{k+1})$ and to report the occurrences in $\mathcal{O}(3^k m^{k+1} + \mbox{\it ed})$ time, where $m$ is the length of the pattern and where {\it ed} the number of matching edit scripts. These bounds are independent of alphabet size. The construction of the structure has time bound by $\mathcal{O}(kN|\Sigma|)$, where $N$ is the number of nodes in the index and $|\Sigma|$ the alphabet size.

\textbf{Keywords:} string algorithms, suffix trees, approximate text matching, text indexing.