### 2006-07-04 paper/main.tex

\documentclass{article}
\usepackage{a4wide}
\usepackage[vlined,linesnumbered,ruled]{algorithm2e}
\input{figures/transfig}
\usepackage[]{graphicx}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{subfigure}
%\usepackage{eepic}

\newcommand*{\CC}{C\nolinebreak\hspace{-.05em}\raisebox{.4ex}{\tiny\bf +}\nolinebreak\hspace{-.10em}\raisebox{.4ex}{\tiny\bf +}}
\newcommand*{\Assign}{\ensuremath\leftarrow}
\newcommand*{\putstring}[1]{\textsl{#1}}
\newcommand*{\operation}[1]{\mbox{\textbf{#1}}}
\newcommand*{\Node}{\ensuremath{\overline{w}}}
\newcommand*{\bigO}{\mathcal{O}}
\graphicspath{{figures/}{graphics/}}
\newtheorem{lemma}{Lemma}
\newtheorem{definition}{Definition}

\title{Dotted Suffix Trees\\A Structure for Approximate Text Indexing}
\author{Lu\'{\i}s Pedro Coelho \and Arlindo L. Oliveira}
%\institute{INESC-ID/IST\\
%\email{\{luis,aml\}@algos.inesc-id.pt}}
\date{}

\begin{document}
\maketitle

\begin{abstract}
\input{abstract}
\end{abstract}

\section{Introduction}\label{sec:introduction}
\input{introduction}
\section{The Indexing Structure}\label{sec:structure}
\input{structure}
\section{Searching}\label{sec:search}
\input{search}
\section{Constructing the Dotted Tree}\label{sec:construction}
\input{construction}
\section{Space Considerations}\label{sec:space}
\input{space}
\section{Experimental Results}\label{sec:experimental}
\input{experimental}
\section{Conclusions}\label{sec:conclusion}
\input{conclusion}
\bibliographystyle{alpha}
\bibliography{biblio}
%\clearpage
%\section*{Appendix}
%\clearpage
%\appendix
%\section{Experimental Results (continued)}\label{appx:experimental}
%
%\input{appendices}
\end{document}